---
title: "Japan international trade"
author: "Mitsuo Shiota"
date: "2019-05-30"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Updated: `r Sys.Date()`

## Summary

https://mitsuoxv.shinyapps.io/jp-trade/

I download Japan international trade monthly data from January 2009, make them tidy, and save them to use for Shiny app. And I have created the Shiny app above.

```{r libraries, include=FALSE}
library(tidyverse)

```
## Function

[e-Stat](https://www.e-stat.go.jp/en) provides summarized data as csv files in its [export page](https://www.e-stat.go.jp/stat-search/files?page=1&layout=datalist&toukei=00350300&bunya_l=16&tstat=000001013137&cycle=1&tclass1=000001013260&tclass2=000001013261&second2=1) and [import page](https://www.e-stat.go.jp/stat-search/files?page=1&layout=datalist&toukei=00350300&bunya_l=16&tstat=000001013137&cycle=1&tclass1=000001013260&tclass2=000001013262&second2=1).

The format of a csv file is a vertically stacked list of 12 monthly plus one annual data, whose rows are areas, columns are goods, and numbers are thousand yen. The number of areas of export changes from 51 to 53 in April 2017. I have made read_url_year function to download a csv file, read it, delete non-data rows, add month column, delete annual data, and make it tidy.

```{r function_read_url_year}
read_url_year <- function(url, year, n_area, start_month = 1) {
  # download
  tf <- tempfile(fileext = ".csv")
  
  httr::GET(url, httr::write_disk(tf))
  
  # read
  data_txt <- readLines(tf)
  
  if (str_detect(data_txt[1], ",")) {
    data <- read_csv(data_txt, skip = 1)
  } else {
    data_csv <- data_txt[str_detect(data_txt, ",")]
    data <- read_csv(data_csv)
  }

  # delete non-data rows
  data <- data %>% 
    filter(`Exp or Imp` %in% c("1", "2"))
  
  # add month column
  months <- vector("character", length = nrow(data))
  
  for (i in 1:(14 - start_month)) {
    if (i * n_area > length(months)) break
    
    months[((i - 1) * n_area + 1):(i * n_area)] <- str_c("0", (i + start_month - 1)) %>% str_sub(start = -2L)
  }
  
  data$month <- months
  
  # delete annual data, and make it tidy
  data <- data %>% 
    filter(month != "13") %>% 
    mutate(month = str_c(year, "-", month, "-01") %>% as.Date()) %>% 
    gather(key = "goods", value = "value", -1, -2, -month)
  
  data$value <- as.numeric(data$value)
  
  data
}

```

## Download

I have to know statInfId number to download through REST API https://www.e-stat.go.jp/stat-search/file-download. I have tried estat_getDataCatalog function in [estatapi package](https://cran.rstudio.com/web/packages/estatapi/estatapi.pdf), but failed. So I manually get them from CSV mark in [each page](https://www.e-stat.go.jp/stat-search/files?page=1&layout=datalist&toukei=00350300&bunya_l=16&tstat=000001013137&cycle=1&year=20190&month=12040604&tclass1=000001013260&tclass2=000001013261&result_back=1&second2=1). I have found statInfId numbers in import are one plus those in export.

After I download data, I have combined them into one data frame, and transformed values from thousand to billion yen.

```{r download, include=FALSE, cache=FALSE}
# Export
# Download 2018-2020
export2020 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031914422&fileKind=1",
                          2020, n_area = 53) %>% 
  mutate(`Exp or Imp` = as.character(`Exp or Imp`)) # only for January

export2019 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031907373&fileKind=1",
                          2019, n_area = 53)

export2018 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031792595&fileKind=1",
                          2018, n_area = 53)

# Download 2017
# December 2017 data don't include January to March, due to n_area change
# Download March 2017 to get January to March
export2017_1_3 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031571360&fileKind=1",
                          2017, n_area = 51)

export2017_4_12 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031670759&fileKind=1",
                              2017, n_area = 53, start_month = 4)

# Download 2009-2016
statid_ex <- tibble(
  year = 2009:2016,
  statInfId = c("000007064353", "000008622970", "000012679430", "000018862219",
                "000023618985", "000028157046", "000031362244", "000031528717")
)

export <- vector("list", length = nrow(statid_ex))

for (i in seq_along(export)) {
  url <- str_c("https://www.e-stat.go.jp/stat-search/file-download?statInfId=", statid_ex$statInfId[i], "&fileKind=1")
  
  export[[i]] <- read_url_year(url, statid_ex$year[i], n_area = 51)
}

# Combine
export_df <- bind_rows(export)

export_df <- export_df %>% 
  bind_rows(export2017_1_3, export2017_4_12,
            export2018, export2019, export2020)


# Import
# Download 2018-2020
import2020 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031914423&fileKind=1",
                            2020, n_area = 51) %>% 
  mutate(`Exp or Imp` = as.character(`Exp or Imp`)) # only for January

import2019 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031907374&fileKind=1",
                            2019, n_area = 51)

import2018 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031792596&fileKind=1",
                            2018, n_area = 51)

# Download 2017
# Contrary to export, n_areas does not change in import, still
# December 2017 data don't include January to March
# Download March 2017 to get January to March
import2017_1_3 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031571361&fileKind=1",
                                2017, n_area = 51)

import2017_4_12 <- read_url_year("https://www.e-stat.go.jp/stat-search/file-download?statInfId=000031670760&fileKind=1",
                                 2017, n_area = 51, start_month = 4)

# Download 2009-2016
statid_im <- tibble(
  year = 2009:2016,
  statInfId = c("000007064354", "000008622971", "000012679431", "000018862220",
                "000023618986", "000028157047", "000031362245", "000031528718")
)

import <- vector("list", length = nrow(statid_im))

for (i in seq_along(import)) {
  url <- str_c("https://www.e-stat.go.jp/stat-search/file-download?statInfId=", statid_im$statInfId[i], "&fileKind=1")
  
  import[[i]] <- read_url_year(url, statid_im$year[i], n_area = 51)
}

# Combine
import_df <- bind_rows(import)

import_df <- import_df %>% 
  bind_rows(import2017_1_3, import2017_4_12,
            import2018, import2019, import2020)


# combine export and import
trade <- export_df %>% 
  bind_rows(import_df)

# from K yen to B yen
trade <- trade %>% 
  mutate(value = value / 1000000)

```

Added areas of export in April 2017 are:

```{r setdiff, echo=FALSE}
setdiff(export2017_4_12$Area %>% unique(),
        export2017_1_3$Area %>% unique())

```

## Charts

I have drawn charts of export to China, as an example.

```{r charts, warning=FALSE, echo=FALSE}
export_china <- trade %>% 
  filter(`Exp or Imp` == 1) %>% 
  filter(Area == "CHINA", goods == "Grand Total")

export_china %>% 
  ggplot(aes(x = month, y = value)) +
  geom_line() +
  labs(
    title = str_c("Export of ", export_china$goods[1], " to ", export_china$Area[1]),
    x = "", y = "billion yen per month"
  )

export_china %>% 
  group_by(goods) %>% 
  arrange(month) %>% 
  mutate(
    lag_12 = lag(value, 12),
    gr = ((value / lag_12) - 1) * 100
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x = month, y = gr)) +
  geom_hline(yintercept = 0, color = "white", size = 2) +
  geom_line() +
  labs(
    title = str_c("Growth rates of export of ", export_china$goods[1], " to ", export_china$Area[1]),
    x = "", y = "YoY percent"
  )

```

## Save data in a rdata file

I save necessary data in a rdata file to use for Shiny app.

```{r save}
# factorize `Exp or Imp`
levels_exim <- trade$`Exp or Imp` %>% unique()

trade$`Exp or Imp` <- factor(trade$`Exp or Imp`, levels = levels_exim)
levels(trade$`Exp or Imp`) <- c("export", "import")

# create menus
levels_area <- trade$Area %>% unique()

levels_goods <- trade$goods %>% unique()

# add gr (growth rates on year-over-year basis)
trade <- trade %>% 
  group_by(`Exp or Imp`, Area, goods) %>% 
  arrange(month) %>% 
  mutate(
    lag_12 = lag(value, 12),
    gr = ((value / lag_12) - 1) * 100
  ) %>% 
  ungroup() %>% 
  select(-lag_12)

# annual data
trade_year <- trade %>% 
  mutate(
    year = lubridate::year(month)
  ) %>% 
  group_by(`Exp or Imp`, Area, goods, year) %>% 
  summarize(value = sum(value)) %>% 
  ungroup()

save(trade, trade_year, levels_area, levels_goods,
     file = "data/jptrade.rdata")

```


EOL